Diamonds - predictive model
========================================================
- Use EDA to develop a quantitative understanding of the diamond market. 
- Build a predictive model of diamonds to figure out whether a given diamond is a good deal or a rip-off
- Exploring this can help understand what goes into the price of a diamond
- Diamonds gave rise to the mining industry in South Africa - world's most advanced economy in the region

### Use a linear regression model to predict diamond price using other variables in the diamonds dataset.

### Scatterplot Review

Let's start by examining two variables in the data set. The scatterplot is a powerful tool to help you understand the relationship between two continuous variables.

 We can quickly see if the relationship is linear or not. In this case, we can use a variety of diamond characteristics to help us figure out whether the price advertised for any given diamond is reasonable or a rip-off.

Let's consider the price of a diamond and it's carat weight. Create a scatterplot of price (y) vs carat weight (x).

#### Limit the x-axis and y-axis to omit the top 1% of values.

```{r Scatterplot Review}
library(ggplot2)
data("diamonds")
names(diamonds)
```


### Range of price

```{r}
summary(diamonds$price)
```

### Trim the top 1% off of the x axis and y axis

```{r}
ggplot(aes(x= carat, y= price), data = diamonds)+ geom_point(fill=I('#F79420'), color=I('black'),shape=21)+xlim(0, quantile(diamonds$carat,0.99))+ylim(0,quantile(diamonds$price, 0.99))
```


### Solution using qplot
```{r}
qplot(data= diamonds, x=carat, y=price, xlim=c(0,quantile(diamonds$carat,0.99)), ylim=c(0,quantile(diamonds$price,0.99)))+geom_point(fill = I('#F79420'), color=I('black'), shape=21)
```

### Solution using ggplot

```{r}
ggplot(aes(x=carat, y=price), data=diamonds)+scale_x_continuous(lim=c(0,quantile(diamonds$carat,0.99)))+scale_y_continuous(lim=c(0,quantile(diamonds$price,0.99)))+geom_point(fill = I('#F79420'), color=I('black'), shape=21)
```

***

### Price and Carat Relationship
- Non linear relationship - exp or something else
- Dispersion/varaiance of the relationship inc as carat inc
- Can add a linear trendline using the stat_smooth with method = 'lm'

```{r}
ggplot(aes(x=carat, y=price), data=diamonds)+scale_x_continuous(lim=c(0,quantile(diamonds$carat,0.99)))+scale_y_continuous(lim=c(0,quantile(diamonds$price,0.99)))+geom_point(color='#F79420', alpha=1/4)+ stat_smooth(method='lm')
```
The linear trendline does not go through the center of the data in some key places - misses curving in the center of the plot and should slope up more towards the end

If we try to make predictions using this we might be off at some places inside and outside of the existing data that we have displayed 


***

### Frances Gerety
#### The diamonds dataset
- Ships with ggplot2
- Contains prices and specs for >50000 diamonds collected in 2008 from diamondsc.info
- Can't just plug in a model no. and look up the price - diamonds are unique

DeBeers Cartel (1888) - Formed to control the global price of diamonds  - 'A diamond is forever'

***

### The Rise of Diamonds
Massive persuasion - diamonds became a status symbol and indispensable in engagement rings 
***

### ggpairs Function
#### First thing you might consider doing is plotting key variables against each other using the ggpairs function 
- You may want to sample first - otherwise the function may take a long time to render the plots
- Also, if there are more than 10 columns there may be too many plotting windows, so subset on columns first if that's the case

Packages

GGally - for this plot
scales - variety of  things
Memisc - summarize the regression
lattice - few other things
MASS - various functions
car - recode variables
reshape - to reshape and wrangle data
dplyr- create interesting summaries and transformations of your data

```{r ggpairs Function}
# install these if necessary
install.packages('GGally')
install.packages('scales')
install.packages('memisc')
install.packages('lattice')
install.packages('MASS')
install.packages('car')
install.packages('reshape')
install.packages('dplyr')


```

```{r}
# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)
library(dplyr)

```


```{r}
# sample 10,000 diamonds from the data set
set.seed(20022012)
# sample 10000 row number in the range 1-length(data) , all columns
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp, axisLabels = 'internal',
        lower = list(continuous = wrap("points", shape = I('.'))), 
        upper = list(combo = wrap("box", outlier.shape = I('.'))))
```

### What are some things you notice in the ggpairs output?

Lower triangle 
- Grouped histograms for qualitative vs qualitative pairs using y as the grouping factor
- Scatterplots for quantitative vs quantitative pairs

Upper triangle
- Grouped histograms for qualitative vs qualitative pairs - this time using x instead of y as the grouping factor
- box plots for qualitative vs quantitative pairs 
- Correlation for quantitative vs quantitative pairs

The "params" argument to the ggpairs function are there to change the shape of the plotted points in the plot matrix, to make them easier to see. GGally 1.0 changes the syntax of these plotting parameters to no longer be part of a params argument, and instead can be specified as follows:

ggpairs(diamond_samp,
  lower = list(continuous = wrap("points", shape = I('.'))),
  upper = list(combo = wrap("box", outlier.shape = I('.'))))
You can click on the packages tab in RStudio to determine which packages have been installed.

You may receive a message when installing the new packages. If so, click cancel, clear your workspace, and try installing the packages again.

In this video, Solomon works with the plyr package. We worked with the dplyr package to manipulate data frames and to create new ones throughout the course. dplyr is the latest version of plyr that is specifically for working with data frames.

Similarly, we worked with the reshape2 package, which is the newest version of the reshape package.

When you duplicate the plot matrix on your local machine, you may want to add a axisLabels = 'internal' argument to your ggpairs function call to have the variable names on the diagonal of the matrix rather than on the outside.

### Look closely at the relationships that correspond to price - what do you notice?

Critical factor driving price is the size/carat weight of the diamonds

Relationship between price and carat is nonlinear. What might explain this pattern?

On the supply side, maybe larger continuous chunks of diamonds without flaws are harder to find than smaller ones - might help explain exponential looking curve

Weight = f (volume) = f(x*y*z)

We might be especially interested in the cube root of carat weight
Leveraging substantive knowledge about your data like this can lead to fruitful transformations

***
### look at price vs carat plot

```{r}
ggplot(aes(x=carat, y=price), data=diamonds)+scale_x_continuous(lim=c(0,quantile(diamonds$carat,0.99)))+scale_y_continuous(lim=c(0,quantile(diamonds$price,0.99)))+geom_point(fill = I('#F79420'), color=I('black'), shape=21)
```

### The Demand of Diamonds

On the demand side , customers in the market for a less expensive smaller diamond are probably more sensitive to price than more well to do customers.
Those buying less than 1 carat would almost surely never buy a diamond if not for the social norm of using a diamond while proposing
Fewer customers can afford a bigger diamond - more than 1 carat - we shouldn't expect the market for bigger diamonds to be as competitive as the one for smaller diamonds 
So it makes sense that the prce and variance increase with carat size

Often the distribution of any monetary variable like dollars will be highly skewed and varyover orders of magnitude - this can be due to path dependence ( rich getting richer) or  multiplicative processes like year on year inflation or a combination of both - so a good idea to look into compressing any such variable by putting it on a log scale

More tips on when to use log transformation

https://www.r-statistics.com/2013/05/log-transformations-for-skewed-and-wide-distributions-from-practical-data-science-with-r/

### Let's examine the distribution of price again  - generate two price histograms

### Create two histograms of the price variable and place them side by side on one output image. We've put some code below to get you started. The first plot should be a histogram of price and the second plot should transform the price variable using log10. Set appropriate bin widths for each plot. ggtitle() will add a title to each histogram.



```{r The Demand of Diamonds}
library(gridExtra)

plot1 <- qplot(x=price, data= diamonds, binwidth=100,fill=I('orange')) + 
  ggtitle('Price')

plot2 <- qplot(x=price, data=diamonds,binwidth=0.01,fill=I('blue')) + scale_x_log10()+
  ggtitle('Price (log10)')

grid.arrange(plot1,plot2)
```

***

### Connecting Demand and Price Distributions
Notes:

***

### Scatterplot Transformation

```{r Scatterplot Transformation}

```


### Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Overplotting Revisited

```{r Sort and Head Tables}

```


```{r Overplotting Revisited}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Other Qualitative Factors
Notes:

***

### Price vs. Carat and Clarity

Alter the code below.
```{r Price vs. Carat and Clarity}
# install and load the RColorBrewer package
install.packages('RColorBrewer')
library(RColorBrewer)

ggplot(aes(x = carat, y = price), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Clarity and Price
Response:

***

### Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut}
ggplot(aes(x = carat, y = price, color = clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Clarity', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Cut and Price
Response:

***

### Price vs. Carat and Color

Alter the code below.
```{r Price vs. Carat and Color}
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = Cut, reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
```

***

### Color and Price
Response:

***

### Linear Models in R
Notes:

Response:

***

### Building the Linear Model
Notes:

```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***

### Model Problems
Video Notes:

Research:
(Take some time to come up with 2-4 problems for the model)
(You should 10-20 min on this)

Response:

***

### A Bigger, Better Data Set
Notes:

```{r A Bigger, Better Data Set}
install.package('bitops')
install.packages('RCurl')
library('bitops')
library('RCurl')

diamondsurl = getBinaryURL("https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda")
load(rawConnection(diamondsurl))
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set
Notes:

```{r Building a Model Using the Big Diamonds Data Set}

```


***

## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

```{r}
#Be sure you’ve loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI.

***

## Final Thoughts
Notes:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

