Lesson 5
========================================================

### Multivariate Data
Two variables - scatterplot works
Third variable - add another way to represent it

***

### Moira Perceived Audience Size Colored by Age
Perceived audience size vs actual audience size
3rd var : age
Perceived audience size by age plot - add color to represent age
- again getting horizontal stripes - no pattern in color - overplotting 
Can't tell if older people guessed better than younger people
Sometimes it's a dead end

***

### Third Qualitative Variable
Examine relationship between friend count and age by adding a third variable

Without adding age as a variable:

```{r}
library(ggplot2)
ggplot(aes(friend_count),
       data = subset(pf, !is.na(gender))) + geom_histogram()+facet_wrap(~gender)
```

Statistics for friedn_count by gender
```{r}
by(pf$friend_count,pf$gender,summary)
```
It appears that females have more friends on average than male users
Is this because female users have a different age distribution??
Maybe conditional on age, the differences are actually larger?

```{r Third Qualitative Variable}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv',sep='\t')
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot()+
  stat_summary(fun.y=mean, geom = 'point', shape =4)
```
Averages are marked by an X
Since male users are a bit younger, we might think that a simple male female comparison doesn't capture the differences in friend count


```{r}
ggplot(aes(x=age, y = friend_count),data = subset(pf, !is.na(gender))) + geom_line(aes(color=gender),stat='summary',fun.y=median)
```

Almost throughout median friend count for women is larger than for men, with exceptions - that include noisy estimates for very "old"" users- we are not confident about these ages reported
Users reporting to be 70 have the same friend count regardless of gender


# Write code to create a new data frame,called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

```{r}
library(dplyr)
sub_pf=subset(pf, !is.na(pf$gender))
pf.fc_by_age_gender <- sub_pf %>%
  group_by(age,gender) %>%
 
  summarise(mean_friend_count = (mean(friend_count)), median_friend_count = round(median(as.numeric(friend_count))), n=n())%>%
   ungroup() 

pf.fc_by_age_gender 
```


***

### Plotting Conditional Summaries
Notes:
Plot median friend count for each gender as age increases
```{r Plotting Conditional Summaries}
ggplot(aes(x=age, y = median_friend_count),data = pf.fc_by_age_gender) + geom_line(aes(color=gender))




```
### Thinking in Ratios
Gender difference is largest for the young users
How many times more friends does the average female user have more than the male user??

***


Notes:

***

### Wide and Long Format
the grouped data frame is in Long format 
It has 2 rows male and female for each age
Convert it to a wide format 
age | male| female with male and female holding the median friend counts

***

### Reshaping Data
Notes:

```{r}
install.packages("tidyr")
library(tidyr)

spread(subset(pf.fc_by_age_gender, 
       select = c('gender', 'age', 'median_friend_count')), 
       gender, median_friend_count)
```
Using reshape2
```{r}
install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, age~ gender, value.var= 'median_friend_count')
head(pf.fc_by_age_gender.wide)

```
 melt - allows conversion from wide to long format
 
 It's important to use quotes around the variable name that is assigned tovalue.var. 

We could also create a similar data frame using the dplyr and tidyr packages:

```{r}
pf.fc_by_age_gender.wide.alt <-
  subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
         !is.na(gender)) %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = male / female)

head(pf.fc_by_age_gender.wide)

```

***

### Ratio Plot

# Plot the ratio of the female to male median friend counts using the data frame
# pf.fc_by_age_gender.wide.

# Think about what geom you should use.Add a horizontal line to the plot with
# a y intercept of 1, which will be the base line. Look up the documentation
# for geom_hline to do that. Use the parameter linetype in geom_hline to make the
# line dashed.

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed 3 = dotted, 4 = dotdash, 5 = longdash 6 = twodash
```{r Ratio Plot}
ggplot(aes(x=age, y= female/male), data=pf.fc_by_age_gender.wide)+geom_line()+geom_hline(yintercept =1, alpha=0.3, linetype = 2)
```
In young female users, median female user has about 2.5 times as many friends as the median male user,
Clearly it was helpful to condition with age in understanding the relationship of friend count with gender
The pattern is robust for users of many different ages - highlights where the difference is striking 

One idea that shows the complexity of our interpretation is - people from particular countries who joined facebook recently are more likely to be male with lower friend counts




***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}

```

***

### Cut a Variable
Notes:

```{r Cut a Variable}

```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}

```

***

### Friending Rate
Notes:

```{r Friending Rate}

```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

