Lesson 5
========================================================

### Multivariate Data
To analyze two variables - scatterplot works
Third variable - add another way to represent it

***

### Moira Perceived Audience Size Colored by Age
Perceived audience size vs actual audience size
3rd var added: age
Perceived audience size by age plot - add color to represent age
- again getting horizontal stripes - no pattern in color - overplotting 
Can't tell if older people guessed better than younger people
Sometimes it's a dead end

***

### Third Qualitative Variable
Examine the relationship between friend count and age by adding a third variable - age

Without adding age as a variable:

```{r}
library(ggplot2)
ggplot(aes(friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_histogram()+
  facet_wrap(~gender)
```

Statistics for friend_count by gender
```{r}
by(pf$friend_count,pf$gender,summary)
```
Comments:
It appears that females have more friends on average than male users
Is this because female users have a different age distribution??
Maybe conditional on age, the differences are actually larger?

```{r Third Qualitative Variable}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv',sep='\t')
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot()+
  stat_summary(fun.y=mean, geom = 'point', shape =4)
```
Averages are marked by an X, because shape =4 
Since male users are a bit younger, we might think that a simple male female comparison doesn't capture the differences in friend count

Let's look at median friend count by age and gender instead:

```{r}
ggplot(aes(x=age, y = friend_count),data = subset(pf, !is.na(gender))) + geom_line(aes(color=gender),stat='summary',fun.y=median)
```

Almost throughout, the median friend count for women is larger than for men, with exceptions that include noisy estimates for very "old" users- we are not confident about these reported ages
Users reporting to be 70 have the same friend count regardless of gender


# Write code to create a new data frame,called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

summarise will remove one layer of grouping when it runs ( the last group - gender)
We need to ungroup one more time to remove the age layer
Arrange the dataframe by age

```{r}
library(dplyr)
#sub_pf=subset(pf, !is.na(pf$gender)) - can use this dataframe instead of filtering

pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age,gender) %>%
  summarise(mean_friend_count = (mean(friend_count)), median_friend_count = round(median(as.numeric(friend_count))), n=n())%>%
   ungroup() %>%
  arrange(age)

pf.fc_by_age_gender 
```


***

### Plotting Conditional Summaries
Notes:
### Construct the plot for the data from pf.fc_by_age_gender
### Plot median friend count for each gender as age increases
```{r Plotting Conditional Summaries}
ggplot(aes(x=age, y = median_friend_count),data = pf.fc_by_age_gender) + geom_line(aes(color=gender))
```
### Thinking in Ratios
Understand more about about how the difference between male and female friend counts varies with age

Gender difference is largest for the young users
How many times more friends does the average female user have than the male user??

***


Notes:

***

### Wide and Long Format
the grouped data frame is in Long format 
- The variable grouped over - male and female has been repeated for each age
It has 2 rows male and female for each age
Convert it to a wide format 
-Should have one row for each age, put the median friend count for male and female in the next two columns
age | male| female with male and female holding the median friend counts

***

### Reshaping Data
Notes:

```{r}
install.packages("tidyr")
library(tidyr)

spread(subset(pf.fc_by_age_gender, 
       select = c('gender', 'age', 'median_friend_count')), 
       gender, median_friend_count)
```
Using reshape2 package
We want  adataframe so we'll use dcast
To get an array or a matrix use acast 

dcast( <dataframe to be changed> , variables you want to keep with a plus sign - (here you want to keep only age) ~ variables whose categories you want as columns (gender) , value.var = <variable with key measurements of the values in the df>)

```{r}
#install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, age~ gender, value.var= 'median_friend_count')
head(pf.fc_by_age_gender.wide)

```
 melt - allows conversion from wide to back to the long format
 
 It's important to use quotes around the variable name that is assigned tovalue.var. 

### We could also create a similar data frame using the dplyr and tidyr packages:

```{r}
pf.fc_by_age_gender.wide.alt <-
  subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
         !is.na(gender)) %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = male / female)

head(pf.fc_by_age_gender.wide)

```

***

### Ratio Plot

# Plot the ratio of the female to male median friend counts vs age
# using the data frame pf.fc_by_age_gender.wide 

# Think about what geom you should use.Add a horizontal line to the plot with
# a y intercept of 1, which will be the base line. Look up the documentation
# for geom_hline to do that. Use the parameter linetype in geom_hline to make the
# line dashed.

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed 3 = dotted, 4 = dotdash, 5 = longdash 6 = twodash
```{r Ratio Plot}
ggplot(aes(x=age, y= female/male), data=pf.fc_by_age_gender.wide)+geom_line()+geom_hline(yintercept =1, alpha=0.3, linetype = 2)
```
In young female users, median female user has about 2.5 times as many friends as the median male user,
Clearly it was helpful to condition with age in understanding the relationship of friend count with gender
The pattern is robust for users of many different ages and highlights where the difference is striking 
there are many processes that can produce this difference - including the biased distribution from which the pseudo facebook data was generated

One idea that shows the complexity of our interpretation is - people from particular countries who joined facebook recently are more likely to be male with lower friend counts




***

### Third Quantitative Variable
We looked at ratio of female to male friend counts across the categorical variable, age
Usually color or shape are the aesthetics to represent such changes over a categorical variable

What if we looked at color or shape over another numerical variable?
For example, friends are accumulated over time - so facebook tenure/days since registering on facebook is important 

Explore friend count, gender, age and tenure - using a 2D display like a scatterplot and bin a quantitative variables and compare the bins
In this case group users by the year that they joined

### Create a variable called year_joined in the pf data frame using the variable tenure and 2014 as the reference year. The variable year joined should contain the year that a user joined facebook.

```{r Third Quantitative Variable}
names(pf)
```
2014 is the current year,  so 2014 - no. of years on facebook 

if 2.2 years is the tenure, 3 should be subtracted from 2014 - ceiling() rounds up to the nearest integer 
floor() rounds down to the the nearest integer

```{r}
pf$year_joined = 2014 - ceiling(pf$tenure/365)
head(pf)

# OR
# pf$year_joined = floor(2014 - (pf$tenure/365))
```
### Testing the floor function- even when decimal is greater than 5 it will round down
```{r}
floor(3.8)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)
```
Value of the year joined are discrete and range is narrow, so let's table the values

```{r}
table(pf$year_joined)
```
This shows the distribution of users in each year joined
Notice that there is not much data about early joiners - To increase the data in each group of the year joined / tenure category 

'Cut' - useful to make discrete variables from continuous/ numerical ones, sometimes with quantile

### Cut the variable year_joined to create 4 buckets of users
# You need to create the following buckets for the
# new variable, year_joined.bucket

#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

# Note that a parenthesis means exclude the year and a
# bracket means include the year.


```{r}

pf$year_joined.bucket <- cut(pf$year_joined,c(2004,2009,2011,2012,2014))
summary(pf$year_joined.bucket)
```
Should include useNA='ifany' , otherwise table() leaves out NA but summary displays NA

```{r}
table(pf$year_joined.bucket,useNA='ifany')
```


***

### Plotting it All Together
Notes:
This was the plot for median friend count vs age where each gender has its own line 
```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

### Create a similar plot of median friend count vs age where each year_joined.bucket has its own line in the graph
# Create a line graph of friend_count vs. age
# so that each year_joined.bucket is a line
# tracking the median user friend_count across
# age. This means you should have four different
# lines on your plot.

# You should subset the data to exclude the users
# whose year_joined.bucket is NA.

```{r}
ggplot(aes(x=age, y=friend_count), data=subset(pf,!is.na(pf$year_joined.bucket)))+geom_line(aes(color=year_joined.bucket),stat='summary', fun.y =median)
```
Comments:

Users with longer tenure tend to have higher friend counts with the exception of old users (80 and above) - red line at the top 
To put  these cohort specific medians in perspective we can change them to cohort specific means and plot the grand mean on here


***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}

```

***

### Friending Rate
Notes:

```{r Friending Rate}

```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

