Lesson 3
========================================================

***
### Pseudo-Facebook User Data
READING DATA:

You could also use the following code to load the Pseudo Facebook data. 
read.delim('pseudo_facebook.tsv') 

The read.delim() function defaults to the tab character as the separator between values and the period as the decimal character. Run ?read.csv or ?read.delim in the console for more details.

```{r}
getwd()
list.files()
pf <- read.csv('pseudo_facebook.tsv',sep='\t')

```

The file has 99000 rows/obs (users) and 15 variables 

***

DISPLAY THE VARIABLES
Goal is to understand user behavior and demographics - what they do on FB and what they use

```{r Pseudo-Facebook User Data}
names(pf)
```

***

### Histogram of Users' Birthdays - DOB by day - no. of users with birthdays on a certain day


```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
names(pf)
qplot(x=dob_day,data=pf) +
  scale_x_continuous(breaks=1:31)

install.packages('ggthemes', dependencies = TRUE)
library(ggthemes)


  

```
```{r}
theme_set(theme_minimal(14))
ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks=1:31)
```

***

#### What are some things that you notice about this histogram?
Response:
One the first day, huge number of people have their birthdays - unusual
On the 31st, fewer number of people - makes sense because not all months have 31st

***

### Moira's Investigation
Notes:
Mismatch between perceived audience size for a post and actual audience (counted)
People underestimated their audience size - typically quarter of actual audience
In ggplot, she plotted a histogram of people's guesses for perceived audience
***


#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response: Birthday post for a friend, 250 friends saw it, 50% of friends (500)


### Estimating audience size
### Perceived Audience Size
Notes:

***
### Faceting
The prev histogram was for no of people vs day of month born
break this into 12 histograms - one for each month

Add a layer called facet_wrap - has a formula  (~ <var that data is split over=dob_month>,no of columns for graphs to display=3 )
```{r Faceting}
qplot(x=dob_day,data=pf) +
  scale_x_continuous(breaks=1:31) +
  facet_wrap(~dob_month,ncol=3)

```
```{r}

```
Note: The use of the scale_x_discrete() layer as shown in the video is deprecated as of ggplot2 version 2.0. You can use scale_x_continuous() instead, or use ggplot() syntax as shown below.

It's best to create the plot in RStudio on your computer and examine at full screen. Faceting in ggplot2

Equivalent ggplot syntax:

```{r}
ggplot(data = pf, aes(x = dob_day)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month)
```


### Facet_wrap
facet_wrap(formula)
facet_wrap(~variable)
same type of plot for each level of the categorical variable  - in this case, dob_day plot, one for each month

### facet_grid
Similar to facet_wrap - facet_grid(formula) - useful when you are passing 2 or more variables
facet_grid( variables to split vertically ~ variables to split horizontally)


#### Let us take another look at our plot. What stands out to you here?
Huge spike on Jan 1 - could be because deafult DOB on FB is Jan 1, or users to protect privacy set up jan 1 as DOB or it's the first choice in the dropdown menu - make these considerations in the context of the data - look out for these types of anomalies

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}

```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}

```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}

```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}

```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}

```

***

### User Ages
Notes:

```{r User Ages}

```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}

```

***

### Likes on the Web
Notes:

```{r Likes on the Web}

```


***

### Box Plots
Notes:

```{r Box Plots}

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!