Lesson 3
========================================================

***
### Pseudo-Facebook User Data
READING DATA:

You could also use the following code to load the Pseudo Facebook data. 
read.delim('pseudo_facebook.tsv') 

The read.delim() function defaults to the tab character as the separator between values and the period as the decimal character. Run ?read.csv or ?read.delim in the console for more details.

```{r}
getwd()
list.files()
pf <- read.csv('pseudo_facebook.tsv',sep='\t')

```

The file has 99000 rows/obs (users) and 15 variables 

***

DISPLAY THE VARIABLES
Goal is to understand user behavior and demographics - what they do on FB and what they use

```{r Pseudo-Facebook User Data}
names(pf)
```

***

### Histogram of Users' Birthdays - DOB by day - no. of users with birthdays on a certain day


```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
names(pf)
qplot(x=dob_day,data=pf) +
  scale_x_continuous(breaks=1:31)

install.packages('ggthemes', dependencies = TRUE)
library(ggthemes)


  

```
```{r}
theme_set(theme_minimal(14))
ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks=1:31)
```

***

#### What are some things that you notice about this histogram?
Response:
One the first day, huge number of people have their birthdays - unusual
On the 31st, fewer number of people - makes sense because not all months have 31st

***

### Moira's Investigation
Notes:
Mismatch between perceived audience size for a post and actual audience (counted)
People underestimated their audience size - typically quarter of actual audience
In ggplot, she plotted a histogram of people's guesses for perceived audience
***


#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response: Birthday post for a friend, 250 friends saw it, 50% of friends (500)


### Estimating audience size
### Perceived Audience Size
Notes:

***
### Faceting
The prev histogram was for no of people vs day of month born
break this into 12 histograms - one for each month

Add a layer called facet_wrap - has a formula  (~ <var that data is split over=dob_month>,no of columns for graphs to display=3 )
```{r Faceting}
qplot(x=dob_day,data=pf) +
  scale_x_continuous(breaks=1:31) +
  facet_wrap(~dob_month,ncol=3)

```
```{r}

```
Note: The use of the scale_x_discrete() layer as shown in the video is deprecated as of ggplot2 version 2.0. You can use scale_x_continuous() instead, or use ggplot() syntax as shown below.

It's best to create the plot in RStudio on your computer and examine at full screen. Faceting in ggplot2

Equivalent ggplot syntax:

```{r}
ggplot(data = pf, aes(x = dob_day)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month)
```


### Facet_wrap
facet_wrap(formula)
facet_wrap(~variable)
same type of plot for each level of the categorical variable  - in this case, dob_day plot, one for each month

### facet_grid
Similar to facet_wrap - facet_grid(formula) - useful when you are passing 2 or more variables
facet_grid( variables to split vertically ~ variables to split horizontally)


#### Let us take another look at our plot. What stands out to you here?
Huge spike on Jan 1 - could be because deafult DOB on FB is Jan 1, or users to protect privacy set up jan 1 as DOB or it's the first choice in the dropdown menu - make these considerations in the context of the data - look out for these types of anomalies

***

### Be Skeptical - Outliers and Anomalies
Types of anomalies/outliers can influence how you deal
-Accurate data about extreme case - someone tweeting excessively
-Bad data/limitations of the data, eg. normal value replaced with an extreme value - for eg, people with high incomes have their value replaced - top coded 

A top-coded data set is one for which the value of variables above an upper bound are censored. This is often done to preserve the anonymity of people participating in the survey.

***

### Moira's Outlier
Most people guessed a small number but atleast one person gave a really huge number (10 million)  - small bar at the end  distant from the initial tall bar - adjust the axis to see the bulk of the data - cut the outlier out 
#### Which case do you think applies to Moiras outlier?
Bad data about an extreme case ( max friends possible is 5000 - 10 million is bad, extreme) Note that this study only asked people about posts with "Friends" privacy, not "Public" privacy settings
Think about what case your outlier falls in - then you know whether to replace / exclude them if you want to make those decisions

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x=friend_count,data=pf)
```

#### How is this plot similar to Moira's first plot?
Data is squished on the left side of the plot like Moira's
Long-tailed data - common for user level data - most users have friend counts under 500 
Max is 5000 - tail extends till there

### Equivalent Code in ggplot for histogram of friend counts
```{r}

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram()
```


***

### Limiting the Axes
Add a parameter inside qplot (xlim - takes a vector start and end position of axis)
Examine friend counts less than 1000

```{r Limiting the Axes}
qplot(x=friend_count,data=pf,xlim = c(0,1000))
```
### Limit axis without xlim

```{r}
qplot(x=friend_count,data=pf) +
  scale_x_continuous(limits=c(0,1000))
```

Equivalent ggplot syntax:


```{r}
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000))
```


Similar function to limit y axis is scale_y_continuos
ggplot allows you to build up your plot in layers

### Exploring with Bin Width
When you adjust the bin width, you see dramatic changes, bin width to 1, taaaalll vertical lines for each number - 10,20,50,100 - don't tend to guess numbers in between 
In most cases, it was quarter the size of the actual audience

***

### Adjusting the Bin Width
Bin width is the width of each bar 
The prev histogram we created is skewed, let's adjust bin width for it - pass a binwidth parameter to qplot and break up the x axis every 50 units - pass breaks 
parameter = sequence( start , end, interval step)
```{r}
qplot(x=friend_count, data=pf, binwidth=25) +
  scale_x_continuous(limits=c(0,1000), breaks=seq(0,1000,50))
```
Taking a closer look, easy to see that many users have less than 25 friends - probably new users

Equivalent ggplot syntax:
```{r}
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```


### Which gender group on average has more friends?
To answer this we need to split the above histogram into 2 plots - males and females
```{r}
library(ggplot2)
```

### Faceting Friend Count histogram by gender
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender,ncol=2)
```
In the alternate solution below, the period or dot in the formula for facet_grid() represents all of the other variables in the data set. Essentially, this notation splits up the data by gender and produces three histograms, each having their own row.
```{r}
qplot(x = friend_count, data = pf) +
  facet_grid(gender ~ .)
```

Equivalent ggplot syntax:

```{r}
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```
Third histogram for NA - this is how R deals with missing values- takes on a value of NA
***

### Omitting NA Values
https://www.statmethods.net/input/missingdata.html
Facet the friend counts histogram by gender ignoring the NA 
use the subset command - first parameter just the dataset, condition is the second parameter
!is.na(gender) - removes rows that have na for the gender variable
! is called bang

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf,!is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender,ncol=2)
```

To omit observations that have any na in them, not just gender - be careful while using this

```{r}
qplot(x = friend_count, data = na.omit(pf), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender,ncol=2)
```
Equivalent ggplot syntax:

```{r}
ggplot(aes(x = friend_count), data = subset(pf, !is.na(gender))) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender

Instead of histogram let's use table
```{r Statistics \'by\' Gender}
table(pf$gender)
```
It looks like slightly more males than females in the sample
Average friend count by gender

by(variable, a categorical var or a list of indices to subset over, a function)

In our case friend count is the variable, gender is the categorical var, and summary is the function to get basic statistics

```{r}
by(pf$friend_count,pf$gender,summary)
```

#### Who on average has more friends: men or women?
Women

#### What's the difference between the median friend count for women and men?
22

#### Why would the median be a better measure than the mean?

Mean friend count is higher than median for males and females - this makes sense since our data is long tailed - values in the higher friend counts will pull the mean to the right (few people will higher friend count drag the mean right) - mean is not representative of most users,  median is a more robust statistic, resistant to change since it marks the halfway point for the data points - since we can trust half our values we can report a reliable location of the center of our distribution

***

### Tenure / How many days someone has been using facebook
Use color to jazz up the plot.The parameter color determines the color outline of objects in a plot.

The parameter fill determines the color of the area inside objects in a plot.

You might notice how the color black and the hex code color of #099DD9 (a shade of blue) are wrapped inside of I(). The I() functions stand for 'as is' and tells qplot to use them as colors.

```{r Tenure}
library(ggplot2)

qplot( x = tenure, data = pf, binwidth =30, color = I('black'), fill = I('#099DD9'))
```
Equivalent ggplot syntax for plots:

```{r}
ggplot(aes(x = tenure), data = pf) +
  geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')

ggplot(aes(x = tenure/365), data = pf) +
  geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')
```

***

#### How would you create a histogram of tenure by year?
Limit to 7 years, change color, change bin width, change x axis to increment by 1 year

```{r Tenure Histogram by Year}
qplot( x = tenure/365, data = pf, binwidth =0.25, color = I('black'), fill = I('#F79420')) +
  scale_x_continuous(breaks=seq(1,7,1),limits=c(0,7))
```

***

### Labeling Plots
By default ggplot2 uses variable names as labels


```{r Labeling Plots}
qplot( x = tenure/365, data = pf, binwidth =0.25, color = I('black'), fill = I('#F79420'), xlab = 'Number of years using Facebook', ylab = 'Number of users in sample') +
  scale_x_continuous(breaks=seq(1,7,1),limits=c(0,7))
```
Equivalent ggplot syntax:
```{r}
ggplot(aes(x = tenure / 365), data = pf) +
  geom_histogram(color = 'black', fill = '#F79420') +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')
```

***

### User Ages - create a histogram of facebook users based on age, play with binwidth, breaks and labels
Notes:
```{r}
summary(pf$age)
```


```{r User Ages}
qplot( x = age, data = pf, binwidth =1, color = I('black'), fill = I('#66FF33'), xlab = 'Age of facebook users (years) ', ylab = 'Number of users in sample') +
  scale_x_continuous(breaks=seq(1,115,5),limits=c(0,115))
```
Equivalent ggplot syntax:

```{r}
ggplot(aes(x = age), data = pf) +
  geom_histogram(binwidth = 1, fill = '#5760AB') +
  scale_x_continuous(breaks = seq(0, 113, 5))
```

#### What do you notice?
Binwidth of 1 helps us spot spikes in the data
less than 13, no users makes sense because users must be atleast 13 years old to use facebook
Unusual spikes above 100 - exaggerating age
***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}

```

***

### Likes on the Web
Notes:

```{r Likes on the Web}

```


***

### Box Plots
Notes:

```{r Box Plots}

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!