LOAD THE DATASET
```{r}
library(ggplot2) #must load the ggplot package first
data(diamonds) #loads the diamonds data set since it comes with the ggplot package
summary(diamonds)
```


## No of observations = 53940
## No of variables = 10
## No of ordered factors = 3

```{r}
# Get details on the dataset
?diamonds
```

## Letter representing the best color for a diamond : D

### Create a histogram of the price of all the diamonds in the diamond dataset
```{r}
qplot(x=price,data=diamonds)
```
The distribution is long tailed
```{r}
summary(diamonds$price)
```
## How many diamonds cost less than $500?

```{r}
sum (diamonds$price < 500)
```

## How many diamondscost less than $250?
```{r}
sum (diamonds$price < 250)
```
## How many diamonds cost $15000 or more?

```{r}
sum (diamonds$price >= 15000)
```

# Explore the largest peak in the
# price histogram you created earlier.

# Try limiting the x-axis, altering the bin width,
# and setting different breaks on the x-axis.

# There won't be a solution video for this
# question so go to the discussions to
# share your thoughts and discover
# what other people find.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Submit your final code when you are ready.


```{r}
qplot(x=price, data=diamonds, binwidth=1) +
  scale_x_continuous(limits=c(250,2000),breaks=seq(250,2000,250))
  
```
No diamonds cost around $1500 - gap in the histogram

Limiting the axes to explore the peak

```{r}
qplot(x=price, data=diamonds, binwidth=1) +
  scale_x_continuous(limits=c(250,700),breaks=seq(250,700,25))
```

```{r}
qplot(x=price, data=diamonds, binwidth=1) +
  scale_x_continuous(limits=c(600,700),breaks=seq(600,700,5))
```

The mode is at $605 (binwidth =1) - most diamonds have this price - largest peak in the histogram

# Break out the histogram of diamond prices by cut.
# You should have five histograms in separate
# panels on your resulting plot.
```{r}
names(diamonds)
```


```{r}
qplot(x=price,data=diamonds) + 
  facet_wrap(~cut)
```
```{r}
by(diamonds$price,diamonds$cut,summary)
```

## Which cut has the highest priced diamond
premium

## Which cut has the lowest priced diamond
Premium and Ideal

## Which cut has the lowest median price?
Ideal

### Scales and multiple histograms
```{r}
qplot(x = price, data = diamonds) + facet_wrap(~cut)
```

# In the last exercise, we looked at the summary statistics
# for diamond price by cut. If we look at the output table, the
# the median and quartiles are reasonably close to each other.

# diamonds$cut: Fair
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     337    2050    3282    4359    5206   18570 
# ------------------------------------------------------------------------ 
# diamonds$cut: Good
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     327    1145    3050    3929    5028   18790 
# ------------------------------------------------------------------------ 
# diamonds$cut: Very Good
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     336     912    2648    3982    5373   18820 
# ------------------------------------------------------------------------ 
# diamonds$cut: Premium
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     326    1046    3185    4584    6296   18820 
# ------------------------------------------------------------------------ 
# diamonds$cut: Ideal
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     326     878    1810    3458    4678   18810 

# This means the distributions should be somewhat similar,
# but the histograms we created don't show that.

# The 'Fair' and 'Good' diamonds appear to have 
# different distributions compared to the better
# cut diamonds. They seem somewhat uniform
# on the left with long tails on the right.

# Let's look in to this more.

# Look up the documentation for facet_wrap in R Studio.
# Then, scroll back up and add a parameter to facet_wrap so that
# the y-axis in the histograms is not fixed. You want the y-axis to
# be different for each histogram.

# If you want a hint, check out the Instructor Notes.

```{r}
qplot(x = price, data = diamonds) + facet_wrap(~cut,scales="free") 
```
Now the scales for y axis are different for different graphs

# Create a histogram of price per carat
# and facet it by cut. You can make adjustments
# to the code from the previous exercise to get
# started.

# Adjust the bin width and transform the scale
# of the x-axis using log10.

```{r}
qplot(x=price/carat, data=diamonds)+
  facet_wrap(~cut,scales="free") +
  scale_x_log10()
```

You find normal-ish distributions and a slightly bimodal distribution for Very Good cuts? 
##(If the Very Good facet looks smooth, try decreasing your bin width! If the data looks skewed, did you remember to perform a log10 transform?)

# Investigate the price of diamonds using box plots,
# numerical summaries, and one of the following categorical
# variables: cut, clarity, or color.

# There won't be a solution video for this
# exercise so go to the discussion thread for either
# BOXPLOTS BY CLARITY, BOXPLOT BY COLOR, or BOXPLOTS BY CUT
# to share you thoughts and to
# see what other people found.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

Price by Cut box plot

```{r}
qplot( x = cut, y = price, data = diamonds, geom = 'boxplot')
```
```{r}
by(diamonds$price,diamonds$cut,summary)
```

Using coord cartesian to limit the y axis and take a closer look at the boxes leaving out the outliers

```{r}
qplot( x = cut, y = price, data = diamonds, geom = 'boxplot') + 
  coord_cartesian(ylim=c(0,7500))
```
Median price for ideal cut is the least while median price for fair cut is the highest

Price by Color Box plots

```{r}
qplot( x = color, y = price, data = diamonds, geom = 'boxplot')
```
Adjusting y axis limit for closer look at the boxes

```{r}
qplot( x = color, y = price, data = diamonds, geom = 'boxplot') +
  coord_cartesian(ylim=c(0,10000))
```
Getting statistics for price by color

```{r}
by(diamonds$price,diamonds$color,summary)
```
Color J has the highest median price while color D has the lowest median price


Price by clarity box plots

```{r}
qplot( x = clarity, y = price, data = diamonds, geom = 'boxplot')
```


Adjusting the y limits to take a closer look at the boxes


```{r}
qplot( x = clarity, y = price, data = diamonds, geom = 'boxplot') +
  coord_cartesian(ylim=c(0,7500))
```
geting statistics

```{r}
by(diamonds$price,diamonds$clarity,summary)
```
Clarity level  SI2 has the highest median price and clarity level IF has the lowest median price


## What is the IQR for diamonds with the best color

```{r}
IQR(subset(diamonds,color=='D')$price)
```

## What is the IQR for diamonds with the worst color?
```{r}
IQR(subset(diamonds,color=='J')$price)
```

Oddity / Trend observed:

The best color,clarity and cut seem to have the lowest median price and worst cut,color and clarity seem to have the highest median price

# Investigate the price per carat of diamonds across
# the different colors of diamonds using boxplots.

# Go to the discussions to
# share your thoughts and to discover
# what other people found.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Copy and paste all of the code that you used for
# your investigation, and submit it when you are ready.




```{r}
qplot( x = color, y = price/carat, data = diamonds, geom = 'boxplot')
```



Adjusting the y axis to take a closer look at the boxes

```{r}
qplot( x = color, y = price/carat, data = diamonds, geom = 'boxplot') + 
  coord_cartesian(ylim=c(0,6000))
```

getting statistics

```{r}
diamonds$pricepercarat <- diamonds$price/diamonds$carat

by(diamonds$pricepercarat,diamonds$color,summary)
```
Color H has the highest median price per carat while color E has the lowest median price per carat

## Investigate the weight of diamonds(carat) using a frequency polygon. Use different bin widths to see how the polygon changes.

## What carat size has a count greater than 2000?? Give all that apply out of the options
```{r}
summary(diamonds$carat)
```

```{r}
qplot(x = carat, data = diamonds ,geom='freqpoly',binwidth=0.02) +
  scale_y_continuous(breaks=seq(2000,30000,1000))+
  scale_x_continuous(limits=c(0.2,5),breaks=seq(0.2,5,0.2))
  
```
```{r}
t=as.data.frame(table(diamonds$carat))
t
```
For the question, we are looking for counts of exact value - try starting out with a larger bin width and making it smaller to see how the plot changes, then focus in on the values of most interest.

Options: 0.1,0.3,0.8,1.01,1.6,2.0,3.0,5.0

Correct ans: 0.3, 1.01
Used table of counts check the values 
***





