Diamond Price Analysis - Two variables

```{r}
getwd()
```
```{r}
library(ggplot2)
data("diamonds")
names(diamonds)
```

### Create a scatterplot of price vs x 

```{r}
ggplot ( aes(x,price), data=diamonds)+geom_point()
```
### Observations
As x increases the price seems to increase exponentially, with some outliers

### Correlation between price and x

```{r}

cor.test(diamonds$x,diamonds$price)
```

###  Correlation between price and y

```{r}
cor.test(diamonds$y,diamonds$price)
```
### Correlation between price and z

```{r}
cor.test(diamonds$z,diamonds$price)
```
### Create scatterplot of price vs depth

```{r}
ggplot ( aes(depth,price), data=diamonds)+geom_point()
```

# Change the code to make the transparency of the
# points to be 1/100 of what they are now and mark
# the x-axis every 2 units.

```{r}
ggplot ( aes(depth,price), data=diamonds)+geom_point(alpha = 1/100)+scale_x_continuous(breaks=seq(0,80,2))
```

### Observations
From the plot we can see that most diamonds are between 59 and 64 values of depth

### Correlation between price and depth
```{r}
cor.test(diamonds$depth, diamonds$price)
```
### Would you use depth to predict price of a diamond, based on the correlation coeff ?

There is no significant variability in price with varying depth - Most diamonds fall in the 59-64 depth bucket, with a considerable range of price 

# Create a scatterplot of price vs carat and omit the top 1% of price and carat values.

```{r}
ggplot ( aes(carat,price), data=diamonds)+geom_point()+xlim(0,quantile(diamonds$carat,0.99)) +
  ylim(0,quantile(diamonds$price,0.99))

```



