Diamond Price Analysis - Two variables

```{r}
getwd()
```
```{r}
library(ggplot2)
data("diamonds")
names(diamonds)
```

### Create a scatterplot of price vs x 

```{r}
ggplot ( aes(x,price), data=diamonds)+geom_point()
```
### Observations
As x increases the price seems to increase exponentially, with some outliers

### Correlation between price and x

```{r}

cor.test(diamonds$x,diamonds$price)
```

###  Correlation between price and y

```{r}
cor.test(diamonds$y,diamonds$price)
```
### Correlation between price and z

```{r}
cor.test(diamonds$z,diamonds$price)
```
### Create scatterplot of price vs depth

```{r}
ggplot ( aes(depth,price), data=diamonds)+geom_point()
```

# Change the code to make the transparency of the
# points to be 1/100 of what they are now and mark
# the x-axis every 2 units.

```{r}
ggplot ( aes(depth,price), data=diamonds)+geom_point(alpha = 1/100)+scale_x_continuous(breaks=seq(0,80,2))
```

### Observations
From the plot we can see that most diamonds are between 59 and 64 values of depth

### Correlation between price and depth
```{r}
cor.test(diamonds$depth, diamonds$price)
```
### Would you use depth to predict price of a diamond, based on the correlation coeff ?

There is no significant variability in price with varying depth - Most diamonds fall in the 59-64 depth bucket, with a considerable range of price 

# Create a scatterplot of price vs carat and omit the top 1% of price and carat values.

```{r}
ggplot ( aes(carat,price), data=diamonds)+geom_point()+xlim(0,quantile(diamonds$carat,0.99)) +
  ylim(0,quantile(diamonds$price,0.99))

```

# Create a scatterplot of price vs. volume (x * y * z).This is a very rough approximation for a diamond's volume.

# Create a new variable for volume in the diamonds data frame. This will be useful in a later exercise.

```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z
ggplot ( aes(volume,price), data=diamonds)+geom_point()
```

Volumes are mostly below 500 with some outliers - steep increase in price as volume increases from 0 to 500

Some volumes are 0, one volume is 4000 and a cheaper diamond with a volume near 900
Depending on your investigation, it may or may not be important for you to understand how outliers, like these, came to be in your data.
```{r}
library(dplyr)
#library(plyr)
#detach("package:plyr", unload=TRUE)
```
### Number of diamonds with volume zero

```{r}
sum(diamonds$volume == 0)
#detach("package:plyr", unload=TRUE)
```
### Correlation of price and volume - exclude diamonds with zero volume or >= 800

```{r}
sub_diamonds=subset(diamonds, (diamonds$volume!=0 & diamonds$volume<800))
cor.test(sub_diamonds$volume, sub_diamonds$price)
```

# Subset the data to exclude diamonds with a volume greater than or equal to 800. Also, exclude diamonds
#with a volume of 0. Adjust the transparency of the  points and add a linear model to the plot.
# Do you think this would be a useful model to estimate the price of diamonds? Why or why not?

```{r}
ggplot ( aes(volume,price), data=sub_diamonds)+geom_point(alpha=1/20)+geom_smooth (method= 'lm', color= 'red')
```

The linear model seems like a reasonable fit

# Use the function dplyr package
# to create a new data frame containing
# info on diamonds by clarity.

# Name the data frame diamondsByClarity

# The data frame should contain the following
# variables in this order.

#       (1) mean_price
#       (2) median_price
#       (3) min_price
#       (4) max_price
#       (5) n

# where n is the number of diamonds in each
# level of clarity.



```{r}
library(dplyr)
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price),
          median_price = median(as.numeric(price)),
          min_price=min(price),
          max_price = max(price) ,
          n=n())
diamondsByClarity
```

### Summary data frames with the mean price by clarity and color.

```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
diamonds_mp_by_clarity 

```



# Your task is to write additional code to create two bar plots
# on one output image using the grid.arrange() function from the package
# gridExtra.


```{r}
library(gridExtra)
```

```{r}
p1 <- ggplot(aes(clarity,mean_price), data=diamonds_mp_by_clarity)+geom_bar(stat='identity',fill='blue')
p2<- ggplot (aes(color,mean_price), data=diamonds_mp_by_color)+geom_bar(stat='identity', fill='pink')
grid.arrange(p1,p2,ncol=1)
```
We think something odd is going here. These trends seem to go against our intuition.


Mean price tends to decrease as clarity improves. The same can be said for color.


We encourage you to look into the mean price across cut.

### Mean price by cut

```{r}
diamondsByCut <- diamonds %>%
  group_by(cut) %>%
  summarise(mean_price = mean(price))
diamondsByCut

```


### Plot bar graph for price by color, clarity and cut
```{r}
p1 <- ggplot(aes(clarity,mean_price), data=diamonds_mp_by_clarity)+geom_bar(stat='identity',fill='blue')
p2<- ggplot (aes(color,mean_price), data=diamonds_mp_by_color)+geom_bar(stat='identity', fill='pink')
p3<- ggplot (aes(cut,mean_price), data=diamondsByCut)+geom_bar(stat='identity', fill='green')

grid.arrange(p1,p2,p3,ncol=1)
```
## Reflections

Premium cut has the highest price while ideal cut has the lowest price

There seems to be some oddity in the data overall as price does not vary from lowest to highest for the worst to best categories for cut, color and clarity of diamonds

