LOAD DATA
```{r}
getwd()
```
```{r}
bdf <- read.csv('birthdaysExample.csv')
names(bdf)
```
## Extract date, month and year from the birthdays
```{r}
library (ggplot2)
#dates_formatted <- as.Date(bdf$dates,format="%m/%d/%y")
#dates_formatted

parsed_dates=strptime(bdf$dates,format="%m/%d/%y")
bdf$date <- strftime(parsed_dates, "%d")
bdf$month <- strftime(parsed_dates, "%m")
bdf$year <- strftime(parsed_dates, "%y")

```

## Histogram of birth date 

```{r}
qplot(x=date, data=bdf)

```
Most common date seems to be the 14th, and 31st has the least number of people born on the date 

## Histogram of birth month

```{r}
qplot(x=month, data=bdf)
```
Most common birth month is March, while the least common seems to be December

## Histogram of birth year

```{r}
table(bdf$year)

```
```{r}
qplot(x=year, data=bdf)
```

It's interesting to observe that every person in this dataset was born in the year 2014
Therefore, in the year 2014, March has been the most common birth month and 14th has been the most common birthdate - which is the pi day!!