
LOAD DATASET

```{r}
library(ggplot2)
data("diamonds")
names(diamonds)
```
# Create a histogram of diamond prices. Facet the histogram by diamond color
# and use cut to color the histogram bars. The plot should look something like this.
# http://i.imgur.com/b5xyrOu.jpg

# Note: In the link, a color palette of type 'qual' was used to color the histogram using
# scale_fill_brewer(type = 'qual').
# Note that the plot in the linked image is using log(price) for the x-axis to adjust for positive skew in the data.

```{r}
ggplot(aes(x=price), data=diamonds)+
  facet_wrap(~color)+
  geom_histogram(aes(fill = diamonds$cut))
```

Adjusting for the positive skew in the data using x=log(price), see how the plot changes

```{r}
ggplot(aes(x=log(price)), data=diamonds)+
  facet_wrap(~color)+
  geom_histogram(aes(fill = diamonds$cut))
```
Using the same colors as in the reference plot


```{r}
ggplot(aes(x=log(price)), data=diamonds)+
  facet_wrap(~color)+
  geom_histogram(aes(fill = diamonds$cut))+
  scale_fill_brewer(type = 'qual')
```


