REDDIT DATA ANALYSIS
-----------------------------------------
Loading reddit data
```{r}
setwd('C:/Users/cdivy/Desktop/Coursework/DAND/Term_2/EDA_using_R/R_practice_code/Reddit_data_analysis')
reddit_df<-read.csv('reddit.csv')
head(reddit_df)
```
```{r}
str(reddit_df)
```
Table the number of people in each emplyment status

```{r}
table(reddit_df$employment.status)
```

Check out the output of summary function - gives tables, statistics

```{r}
summary(reddit_df)
```
Examine the levels of age.range 

```{r}
levels(reddit_df$age.range)
```
Install ggplot2.


```{r}
#install.packages('ggplot2', dependencies = T)
#library(ggplot2)
```

Create a plot to show how many users are in each bin 
```{r}
library(ggplot2)
qplot(data=reddit_df,x=age.range)
```

Can we shift under 18 to the left? ? - Ordered levels / factors
Make more readable plots - 'in come level' can also be converted to an ordered factor

Try a plot for income level
```{r}
qplot(data=reddit_df,x=income.range)
```

The problem is much worse - hard to read. Write code to order the factors - age.range


```{r}
reddit_df$age.range <- ordered(reddit_df$age.range, levels=c("Under 18" ,"18-24","25-34" , "35-44","45-54","55-64","65 or Above"))
qplot(data=reddit_df,x=age.range)

```

Order the levels for income range and create a plot
```{r}
levels(reddit_df$income.range)
```


```{r}
reddit_df$income.range <- ordered(reddit_df$income.range, levels=c("Under $20,000" , "$20,000 - $29,999" , "$30,000 - $39,999"  , "$40,000 - $49,999",  "$50,000 - $69,999", "$70,000 - $99,999" , "$100,000 - $149,999" , "$150,000 or more" ))
qplot(data=reddit_df,x=income.range)
```

