LOAD THE DATASET
```{r}
library(ggplot2) #must load the ggplot package first
data(diamonds) #loads the diamonds data set since it comes with the ggplot package
summary(diamonds)
```


## No of observations = 53940
## No of variables = 10
## No of ordered factors = 3

```{r}
# Get details on the dataset
?diamonds
```

## Letter representing the best color for a diamond : D

### Create a histogram of the price of all the diamonds in the diamond dataset
```{r}
qplot(x=price,data=diamonds)
```
The distribution is long tailed
```{r}
summary(diamonds$price)
```
## How many diamonds cost less than $500?

```{r}
sum (diamonds$price < 500)
```

## How many diamondscost less than $250?
```{r}
sum (diamonds$price < 250)
```
## How many diamonds cost $15000 or more?

```{r}
sum (diamonds$price >= 15000)
```

# Explore the largest peak in the
# price histogram you created earlier.

# Try limiting the x-axis, altering the bin width,
# and setting different breaks on the x-axis.

# There won't be a solution video for this
# question so go to the discussions to
# share your thoughts and discover
# what other people find.

# You can save images by using the ggsave() command.
# ggsave() will save the last plot created.
# For example...
#                  qplot(x = price, data = diamonds)
#                  ggsave('priceHistogram.png')

# ggsave currently recognises the extensions eps/ps, tex (pictex),
# pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

# Submit your final code when you are ready.


```{r}
qplot(x=price, data=diamonds, binwidth=1) +
  scale_x_continuous(limits=c(250,2000),breaks=seq(250,2000,250))
  
```
No diamonds cost around $1500 - gap in the histogram

Limiting the axes to explore the peak

```{r}
qplot(x=price, data=diamonds, binwidth=1) +
  scale_x_continuous(limits=c(250,700),breaks=seq(250,700,25))
```

```{r}
qplot(x=price, data=diamonds, binwidth=1) +
  scale_x_continuous(limits=c(600,700),breaks=seq(600,700,5))
```

The mode is at $605 (binwidth =1) - most diamonds have this price - largest peak in the histogram

